### 2018-08-12 extended_abstract/appendix.tex

%!TEX root = main.tex
\color{red}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_strict_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{strict}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_strict_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_loose_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{loose}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_loose_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_real_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{real}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_real_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_strict_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{strict}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_strict_2}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_loose_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{loose}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_loose_2}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_real_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{real}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_real_2}
\end{figure}