### 2018-08-12 thesis_report/Thesis_Acknowledgements.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Acknowledgments.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified :  May 2018                                        %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\acknowledgments}

% Add entry in the table of contents as section
\addcontentsline{toc}{section}{\acknowledgments}

Firstly I would like to thank my supervisor, Prof. Pedro Adão, for accepting this challenge, as well as his support during this period of time.
I would also like to thank INOV - Inesc Inovação, specifically its cyber unit, who believed in my work and provided means for it to be developed.

Lastly, I would like to thank my family, significant other, friends and colleagues for their support and feedback throughout this time.